%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LatexTemplate
% November 10, 2012
% By Simon Pratt (mostly)
\documentclass[11pt]{article}

\usepackage{Assignment}
\usepackage{CGAlgorithms}
\usepackage{QuestionAnswer}
\usepackage{TheoremStuff}
\usepackage{HeaderStuff}
\usepackage{url}
\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicolumn stuff
\usepackage{multicol}
\usepackage{newclude} % \include without \clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Configuration
\fancyhead[L]{COMP4106 - Final Project}
\fancyhead[R]{Simon Pratt - 100663987}
\bibliographystyle{plain}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document
\begin{document}
\begin{multicols}{2} % remove line if you don't want multiple columns

% Content starts here

\section{The Problem}

The problem domain is to find frequent patterns in large data sets.

\subsection{Motivation}

Frequent patterns in data can give us insight into the semantic
meaning of the data.  In particular, if we are mining for frequent
patterns in a data set of actions we can identify typical workflow and
anomalous activity, which could be useful for security or process
optimization.

\section{Algorithms}

I implemented the following algorithms for mining frequent patterns:
\texttt{Apriori}, \texttt{FP-growth}, and \texttt{Eclat}
\cite{Han2007,fpmlecture}.

\subsection{Apriori}

In which we describe the Apriori algorithm.

\subsection{FP-Growth}

In which we describe the FP-Growth algorithm.

\subsection{Eclat}

In which we describe the Eclat algorithm.

\section{Challenges}

In which we describe the challenges encountered.

\section{Testing Methodology}

In which we describe the methodology of the testing.

\section{Results}
\label{sec:results}

In which we describe the results of the experiments.

\section{Future Work}

One possible direction in which to extend this study is to use the
results of mining datasets for frequent patterns to learn association
rules between items. \cite{wiki_arl}

\appendix
\section{Running Requirements}

To run the code, you simply need to have \texttt{Python 2.6}
installed.  \texttt{Python 2.7} should also work, but the first line
in each python script references \texttt{Python 2.6} in particular,
which should be changed if running on \texttt{Python 2.7}.

There are 4 python scripts included in the code directory:
\texttt{dataset.py}, \texttt{fp\_mining.py}, \texttt{timing.py}, and
\texttt{unit\_tests.py}.

\texttt{dataset.py} simply handles the parsing of files into a simple
datastructure to use for pattern mining.  Running this file with the
name of a datafile as a parameter will run some simple tests on the
module and print the input dataset in its \texttt{Python} format.  The
expected file format is described in the next appendix.

\texttt{fp\_mining.py} containins the algorithm logic for
\texttt{Apriori}, \texttt{FP-Growth}, and \texttt{Eclat}.  The
expected usage of this script is as follows:

\texttt{fp\_mining.py [file] [k] [results]}

Where [file] is the name of an input file, [k] is the desired length
of patterns to mine, and [results] is an optional positive integer
specifying how many results to print.  If [results] is ommitted, all
resulting patterns will be printed.

Note also that this script produces detailed logging output to the
standard error stream on unix, which can be redirected by use of the
\texttt{2>} operator.

\texttt{timing.py} runs the timed tests used to generate the output
discussed in the Results section.  This runs a sequence
of timed tests and prints aggregate information.

\texttt{unit\_tests.py} runs the unit tests for the code base.

\section{Datafile Format}

The datafiles used for testing were retrieved from the FIMI dataset
repository \cite{fimiDatasetRepo}.  The datasets are in a horizontal
format in which each line is an itemset and the items within the set
are integers separated by a single space.  For example, the
hand-written \texttt{tiny.dat} dataset is simply:

\texttt{1 2 3\newline
1 2 4\newline
1 5 6\newline
1 5 7}

The first line is an itemset containing items 1, 2, and 3.  An
important part of the \texttt{Eclat} algorithm is conversion from the
horizontal format to vertical format.

The vertical format of the same data is a dictionary wherein each item
is a key and the value is a \emph{transaction id} set, or
\texttt{tidset}, which contains the IDs of each transaction containing
the key item.

The equivalent vertical format for the \texttt{tiny.dat} dataset is:

\begin{align*}
  1 &: \{ 0, 1, 2, 3 \} \\
  2 &: \{ 0, 1 \} \\
  3 &: \{ 0 \} \\
  4 &: \{ 1 \} \\
  5 &: \{ 2, 3 \} \\
  6 &: \{ 2 \} \\
  7 &: \{ 3 \} \\
\end{align*}

And in \texttt{Python} representation as a dict:

\texttt{\{1:set([0,1,2,3]),2:set([0,1]),3:set([0]),\newline
4:set([1]),5:set([2,3]),6:set([2]),7:set([3])\}}

% Content ends here

\bibliography{references}

\end{multicols} % remove line if you don't want multiple columns
\end{document}
